---
title: "Explore response data!"
output: 
  github_document:
    toc: yes
    toc_depth: 2
---

```{r echo=FALSE, warning=FALSE, message=F}
suppressMessages(library(tidyverse))
suppressMessages(library(xtable))
suppressMessages(library(pheatmap))
suppressMessages(library(RColorBrewer))
suppressMessages(library(ggpubr))
```

```{r echo=F,warning=FALSE, message=FALSE, results='asis'}
load("../data/df.all.RData")
df <- df.all
df <- df %>% filter(timepoint == 'pre')
df <- df %>% merge(., c(Hugo2016 = "Dataset 1", Prat2017 = "Dataset 2", Riaz2017 = "Dataset 3", TCGA = "Dataset 4", VanAllen2015 = "Dataset 5"), by.x = 'dataset', by.y = 'row.names', all.x = T)
colnames(df)[21] <- 'dataset_anon'
d_stats <- df %>% split(.$dataset) %>% map(~ unique(dplyr::select(.,sample, dataset, treatment, dataset_anon))) %>% map(~ cbind(unique(dplyr::select(., dataset, treatment, dataset_anon)), samples = dim(.)[1])) %>% do.call(rbind,.) 
print(xtable(x = d_stats[,-1]),type='html')
```

Scale data by dataset mean and standard deviation

```{r echo=F,warning=FALSE, message=FALSE}
ggplot(df,aes(x=sample,y=scaled_value)) + geom_boxplot(outlier.shape=NA) + theme_classic()
```

```{r echo=F,warning=FALSE, message=FALSE} 
## Uses ggpubr to add pairwise p-values
plot_gene_stats <- function (d,gene_of_interest){
  d %>% filter(Gene==gene_of_interest) %>% dplyr::select(scaled_value,dataset,objective_response) %>% .[complete.cases(.),] %>% 
ggboxplot(x = "dataset", y = "scaled_value",color="objective_response",add = "jitter") + 
    stat_compare_means(aes(group=objective_response),label = "p.format",method = "t.test") +
    ggtitle(gene_of_interest)
}
## Hmm, some package versions (?) do not correctly do the two-factor jitter
## See http://www.sthda.com/english/articles/24-ggpubr-publication-ready-plots/76-add-p-values-and-significance-levels-to-ggplots/
## This should work
## ggboxplot(ToothGrowth, x = "dose", y = "len", color = "supp", palette = "jco",add = "jitter")
```

```{r echo=F,warning=FALSE, message=FALSE} 
## Uses ggpubr to add pairwise p-values
## Anonymize datasets
plot_gene_stats_anon <- function (d,gene_of_interest){
  d %>% filter(Gene==gene_of_interest) %>% dplyr::select(scaled_value,dataset_anon,objective_response) %>% .[complete.cases(.),] %>% 
ggboxplot(x = "dataset_anon", y = "scaled_value", color="objective_response", add = "jitter") + 
    stat_compare_means(aes(group=objective_response), label = "p.format", method = "t.test", 
                       label.y = df %>% filter(Gene==gene_of_interest) %>% dplyr::select(scaled_value,dataset_anon,objective_response) %>% .[complete.cases(.),] %>% .$scaled_value %>% max()*1.5) +
    ggtitle(paste(gene_of_interest, 
                  " (anova, p = ", 
                  d %>% filter(Gene==gene_of_interest) %>% dplyr::select(scaled_value,dataset_anon,objective_response) %>% .[complete.cases(.),] %>% 
                    lm(scaled_value ~ dataset_anon + objective_response, data=.) %>% anova() %>% data.frame() %>% .[2,5] %>% round(5), ")", sep = "")) +
    ylab('Scaled gene expression') + xlab('')
}
## Hmm, some package versions (?) do not correctly do the two-factor jitter
## See http://www.sthda.com/english/articles/24-ggpubr-publication-ready-plots/76-add-p-values-and-significance-levels-to-ggplots/
## This should work
## ggboxplot(ToothGrowth, x = "dose", y = "len", color = "supp", palette = "jco",add = "jitter")
```

## Test gene list for differences in response

### Just plot a couple of genes of interest for sitc presentation

```{r echo=F, warning=FALSE, message=FALSE}
sitc_gois <- c('PDCD1','CD274','CTLA4','HAVCR2','LAG3','CD8A','GZMB','PRF1')
```

Of `r length(sitc_gois)` genes of interest, `r length(intersect(sitc_gois,unique(df$Gene)))` are represented by gene symbol in the IMPRES set.

For each of those, compute ANOVA for linear model, with dataset and objective_response as effects.

```{r echo=F,warning=FALSE, message=FALSE}
d <- df %>% filter(Gene %in% sitc_gois) %>% dplyr::select(Gene,scaled_value,dataset,objective_response) %>% .[complete.cases(.),] 

dd <- d %>% split(.$Gene) %>%
            map( ~ lm(scaled_value ~ dataset + objective_response, data=.)) %>%
            map(anova) %>%
            map(~. ["objective_response","Pr(>F)"]) %>%
            unlist()

```

P-values for the genes of interest
```{r results='asis', echo=FALSE}
a <- as.matrix(sort(dd,decreasing = F));  colnames(a)  <- "P-value"
print(xtable(a,digits=5),type='html')
```

Plots for each gene
```{r echo=F,warning=FALSE, message=FALSE}
for(i in row.names(a)){
  #pdf(paste("../result/",i,".pdf", sep = ""), width = 5, height = 4)
  print(plot_gene_stats_anon(df, i))
  #dev.off()
}
```


### Longer list 

```{r echo=F, warning=FALSE, message=FALSE}
gois_long <- read_tsv(file="../data/curated_gene_list.txt")[[1]]
## 551 of 715 gois have match IMPRES gene symbols
```
Of `r length(gois_long)` genes of interest, `r length(intersect(gois_long,unique(df$Gene)))` are represented by gene symbol in the IMPRES set.

For each of those, compute ANOVA for linear model, with dataset and objective_response as effects.

```{r echo=F,warning=FALSE, message=FALSE}
d <- df %>% filter(Gene %in% gois_long) %>% dplyr::select(Gene,scaled_value,dataset,objective_response) %>% .[complete.cases(.),] 
dd <- d %>% split(.$Gene) %>%
            map( ~ lm(scaled_value ~ dataset + objective_response, data=.)) %>%
            map(anova) %>%
            map(~. ["objective_response","Pr(>F)"]) %>%
            unlist()

```

Lowest p-values

```{r results='asis', echo=FALSE}
a <- as.matrix(head(sort(dd,decreasing = F), 20));  colnames(a)  <- "P-value"
print(xtable(a,digits=5),type='html')
```

Plots for top genes

```{r echo=F, message=FALSE, warning=FALSE, paged.print=TRUE}
for(i in row.names(a)){
  print(plot_gene_stats_anon(df, i))
}
```

### More focused list

```{r echo=F,warning=FALSE, message=FALSE}
gois_short <- read_tsv(file="../data/focused_gene_list.txt")[[1]]
```

Of `r length(gois_short)` genes of interest, `r length(intersect(gois_short,unique(df$Gene)))` are represented by gene symbol in the IMPRES set.

For each of those, compute ANOVA for linear model, with dataset and objective_response as effects.

```{r echo=F,warning=FALSE, message=FALSE}
d <- df %>% filter(Gene %in% gois_short) %>% dplyr::select(Gene,scaled_value,dataset,objective_response) %>% .[complete.cases(.),] 

dd <- d %>% split(.$Gene) %>%
            map( ~ lm(scaled_value ~ dataset + objective_response, data=.)) %>%
            map(anova) %>%
            map(~. ["objective_response","Pr(>F)"]) %>%
            unlist()

```

P-values for the genes of interest
```{r results='asis', echo=FALSE}
a <- as.matrix(sort(dd,decreasing = F));  colnames(a)  <- "P-value"
print(xtable(a,digits=5),type='html')
```

Plots for each gene
```{r echo=F,warning=FALSE, message=FALSE}
for(i in row.names(a)){
  print(plot_gene_stats_anon(df, i))
}
```
